---
title: "Simulating sampling"
output:
  html_document:
    df_print: paged
---

# Simulating probability of sampling during shedding

Assumptions:
- One individual shedding >= 1 day this month
- Average 16 days shedding
- 30 days in 1 month
- Equal probability of sampling on any given day 1-30

```{r setup}

library(tidyverse)
library(magrittr)
library(intervals)
library(patchwork)

# Fix default ggsave background
ggsave <- function(..., bg = 'white') ggplot2::ggsave(..., bg = bg)
theme_set(theme_minimal())

outdir <- here::here("results","sampling_probabilities")

```

## Shedding period

Average duration of infectious shedding of 16 days in an assumed highly vaccinated population.

Modelling study with Isreali data estimated 29 days, assuming population with 0-3 doses of OPV, but only 3 days of high/infectious shedding: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9114981/ 

```{r define_shedding}

Iter = 1000

# Average days of shedding
shed_avg = 16

# Iterations of shed duration
shed_days = rpois(Iter*2, shed_avg)

# Define period of days on which shedding could start in order to coincide with this month (excluding 0)
# per = -(max(shed_days)):30
# per = per[per != 0]

# Random start date of shedding
# shed_start = sample(-16:30, Iter, replace = T)
shed_start = sapply(shed_days, function(x) sample(-60:30, 1, replace = T))
shed_end = shed_start + shed_days

shed_start %>% as.factor() %>% table() %>% barplot()

# png(here::here(outdir, "sim1_shed_start.png"), height = 400, width = 600)
# shed_start %>% as.factor() %>% table() %>% barplot()
# dev.off()

# Define shedding intervals
sims <- data.frame(shed_start, shed_end)

# Keep only draws which intersect with month t=1:30
int_intersect <- function(x){
  tmp <- NA
  for (t in 1:30){
    tmp[t] <- between(t, x[1], x[2])
  }
  if(any(tmp == T)){
    return(TRUE)
  }else{
    return(FALSE)
  }
}

caught <- apply(sims, 1, int_intersect)
sims <- sims[caught,]

# Shed start given intersection with this month
sims$shed_start %>% as.factor() %>% table() %>% barplot()
# 
# png(here::here(outdir, "sim1_shed_start_cond.png"), height = 400, width = 600)
# sims$shed_start %>% as.factor() %>% table() %>% barplot()
# dev.off()

```

-> Approximately uniform for start day of shedding
-> Lower end tails off when conditioning on shedding intersecting with this month

## Sampling

```{r define_sampling}

# Sample taken any day from 1:30
sims$sample_taken = sample(1:30, nrow(sims), replace = T)

hist(sims$sample_taken)

```

## Shedding catchment

```{r simulate_shedding_catchment}

# Sample captures shedding if day of sample is during shedding period
sims %<>% 
  rowwise() %>% 
  mutate(caught = between(sample_taken, shed_start, shed_end)) %>% 
  ungroup() %>% 
  arrange(caught) %>% 
  mutate(i = 1:nrow(sims), tmp = 1)

summary(sims)

```

```{r summary_sims1}

mean(sims$caught)
# 0.36

sims_samp <- slice_sample(sims,n = 100)

ggplot()  +
  geom_hline(aes(yintercept = c(1, 30)))+
  geom_linerange(data = sims_samp,
                 aes(ymin = shed_start, ymax = shed_end, x = i),
                 col = "grey") +
  geom_point(data = sims_samp,
             aes(x = i, y = sample_taken, col = caught)) +
  coord_flip() +
  scale_y_continuous(breaks = c(1,30), labels = c(1,30)) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = NULL, y = NULL, col = "Caught") -> p1

ggplot()  +
  geom_hline(aes(yintercept = c(1, 30)))+
  geom_linerange(data = sims_samp, 
                 aes(ymin = shed_start, ymax = shed_end, x = i, col = caught)) +
  coord_flip() +
  scale_y_continuous(breaks = c(1,30), labels = c(1,30)) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = NULL, y = NULL, col = "Caught") -> p2

p1 + p2
ggsave(here::here(outdir, "sim1_shedding_sampling.png"), width = 12, height = 6)

```

## Simulate for different sampling frequencies

```{r define_fcn}

get_sims <- function(n_samp = 1, Iter = 1000){
  sims2 = data.frame(d = NA_integer_, 
                     start = NA_integer_, 
                     end = NA_integer_, 
                     sample_taken = NA_character_, 
                     caught = NA_character_)
  for (i in 1:Iter){
    
    # Draw shedding duration
    d = rpois(1, shed_avg)
    
    # per = -d:30
    # start = sample(per, 1)
    
    # Draw start day, conditional on start:end intersecting with 1:30
    start <- end <- -99
    while(!int_intersect(c(start, end))){
        start = sample(-60:30, 1)
        end = start + d
    }
    
    # Draw n_samp sampling days 
    samp = sample(1:30, n_samp)
    
    # Flag where sampling days occur during shedding
    caught = any(between(samp,start,end))
    
    # Output
    sims2[i,] = data.frame(d = d, 
                           start = start, end = end,
                           sample_taken = paste0(samp, collapse = ","),
                           caught = (caught == TRUE))
  }
  
  return(sims2)
  
}

sims.list <- lapply(1:4, get_sims, Iter = Iter)

```

```{r plot_sims2}

sims_samp <- slice_sample(sims.list[[1]], n = 100) %>%
  arrange(caught) %>% 
  mutate(sample_taken = as.numeric(sample_taken),
         i = 1:100, tmp = 1)

ggplot()  +
  geom_hline(aes(yintercept = c(1, 30)))+
  geom_linerange(data = sims_samp,
                 aes(ymin = start, ymax = end, x = i),
                 col = "grey") +
  geom_point(data = sims_samp,
             aes(x = i, y = sample_taken, col = caught)) +
  coord_flip() +
  scale_y_continuous(breaks = c(1,30), labels = c(1,30)) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = c(0.1,0.9)) +
  labs(x = NULL, y = NULL, col = "Caught") -> p1

ggplot()  +
  geom_hline(aes(yintercept = c(1, 30)))+
  geom_linerange(data = sims_samp, 
                 aes(ymin = start, ymax = end, x = i, col = caught)) +
  coord_flip() +
  scale_y_continuous(breaks = c(1,30), labels = c(1,30)) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(x = NULL, y = NULL, col = "Caught") -> p2

p1 + p2
ggsave(here::here(outdir, "sim_shedding_sampling.png"), width = 12, height = 6)

```

```{r sum_sims_list}

data.frame(n = 1:4,
           p = sapply(sims.list, function(x) mean(x$caught == TRUE))) -> p_catch_byn

p_catch_byn
# n    p
# 1 0.36
# 2 0.57
# 3 0.66
# 4 0.75

ggplot(p_catch_byn, aes(n, p)) +
  geom_line() +
  labs(x = "Samples per 30 days", y = "Probability",
       title = "Simulated probability of catching shedding of one infectious individual",
       subtitle = paste0("Mean across ",Iter," simulations"))

```

## Simulate sampling distribution of this probability

```{r sim_samp_dist}

get_sims_rep <- function(rep, n_samp = 1, Iter = 1000){
  out <- lapply(1:rep, function(x) get_sims(n_samp = n_samp, Iter = Iter))
  # Return just the proportion of sims in which shedding coincides with sampling
  p <- sapply(out, function(x) mean(x$caught == TRUE))
  return(p)
}

sims_rep_1 <- get_sims_rep(rep = 100, n_samp = 1, Iter = 1000)
hist(sims_rep_1)

png(here::here(outdir, "sim_sampdist_freq1.png"), height = 400, width = 600)
hist(sims_rep_1)
dev.off()

```

```{r sim_samp_dist_byn}

freq <- 1:30
pb <- progress::progress_bar$new(format = "[:bar] :current/:total (:percent)", 
                                   total = length(freq))
sims_rep <- lapply(freq, function(n) {
                            pb$tick()
                            get_sims_rep(rep = 1000, 
                                         n_samp = n, 
                                         Iter = 100)
                          })

saveRDS(sims_rep, here::here(outdir,"sims.rds"))

```

```{r sum_samp_dist_byn}

sims_rep %>% 
  setNames(freq) %>% 
  bind_cols() %>% 
  pivot_longer(everything(), names_to = "SampleFreq", values_to = "p") %>% 
  mutate(SampleFreq = factor(SampleFreq, levels = freq)) -> sum_sims

sum_sims %>% 
  ggplot(aes(x = p, fill = SampleFreq)) +
  geom_histogram(position = "identity", alpha = 0.5) +
  labs(x = "Proportion of sims in which shedding coincides with sampling",
       y = "Frequency",
       fill = "Samples taken")
ggsave(here::here(outdir, "sim_p_catch_hist.png"), height = 5, width = 8)

sum_sims %>% 
  ggplot(aes(x = as.factor(SampleFreq), y = p)) +
  geom_boxplot() +
  labs(x = "Samples taken in 30 days", y = "Estimate",
       # title = "Estimated probabilty of sampling coinciding with shedding",
       # subtitle = "Assuming one infected individual shedding for >= 1 day",
       caption = "Summary over 1000 replicates") -> sim_p_catch_box

sim_p_catch_box
ggsave(here::here(outdir, "sim_p_catch_box.png"), height = 5, width = 8)

p1  + sim_p_catch_box + plot_layout(widths = c(1,2))
ggsave(here::here(outdir, "sim_sampling_prob.png"), width = 12, height = 5)

```

```{r output_samp_dist_byn}

sum_sims %>% 
  group_by(SampleFreq) %>% 
  summarise(EstMean = mean(p),
            EstLwr = quantile(p, 0.025),
            EstUpr = quantile(p, 0.975)) %>% 
  ungroup() -> tab_p

write_csv(tab_p, here::here("inputs", "ES_envsamp_vals.csv"))

ggplot(tab_p, aes(x = SampleFreq,
                  y = EstMean, 
                  ymin = EstLwr, ymax = EstUpr,
                  group = 1)) +
  geom_ribbon(alpha = 0.5, fill = "steelblue") +
  geom_line(col = "steelblue") +
  labs(x = "Samples taken in 30 days", y = "Estimate",
       title = "Estimated probabilty of sampling coinciding with shedding",
       subtitle = "Assuming one infected individual shedding for >= 1 day",
       caption = "95% quantile interval over 1000 replicates")
ggsave(here::here(outdir, "sim_p_catch.png"), height = 5, width = 8)

```

```{r count_days_shedding}

count_days <- data.frame(t = 1:30,
                         count = NA)
for (t in 1:30){
  sum(apply(select(sims.list[[1]],start,end),1, 
            function(x) between(t,x[1],x[2]))) -> count_days$count[t]
}
count_days %>% 
  mutate(dens = count/sum(count)) %>% 
  ggplot(aes(x = t, y = dens)) +
  geom_col()

```
Roughly uniform density of shedding for every day of the month

for each day in the month t = 1:30 what is the probability that:
 + the individual is shedding
 + a sample is taken?

Given an individual sheds for 16 days, there are 45 periods a person can be
shedding during a month of 30 days
There are 16 ways an individual is shedding on any given day 1:30
=> probability person shedding on any given day in a month is 16/45(??)

<!-- freq = 1 # frequency per month -->
<!-- p_detect = data.frame(t = 1:30, -->
<!--                    p_shed = 16/45, -->
<!--                    p_sample = freq*1/30) %>% -->
<!--  mutate(p_shed_sample = p_shed*p_sample) -->
